Interactive Checks for Coordination Avoidance

### 2019-08-18 doc/vldb_slides/sections/bank_example.tex
Slimmed down talk.

  \note{%
    One way to replicate our database is by using coordination to implement
    strong consistency. For example, imagine a user issues a deposit of 10
    dollars to replica A, and simultaneously, a user issues a deposit of 20
    dollars to replica B. The two replicas can coordinate, for example by
    running Paxos, and agree on a single total order in which to execute the
    two transactions. Here, the replicas agree to deposit 10 dollars and then
    deposit 20 dollars. Both transactions succeed, and the replicas send back OK
    messages to the clients letting them know everything went smoothly.
  }
\end{frame}

@@

  \note{%
    Now, assume two clients issue concurrent withdrawal transactions of 30
    dollars and 40 dollars. Again, the two replicas coordinate and agree to
    execute the withdrawal of 30 dollars and then the withdrawal of 40 dollars.

    The withdrawal of 30 dollars executes fine, brining my bank account balance
    to 20 dollars, but the withdrawal of 40 dollars fails because I only have
    20 dollars left. So replica A sends back an OK message to the client, but
    replica B sends back a NO message letting the client know the transaction
    failed. And while the transaction failed, you'll note that we never
    violated our application invariant.
  }
\end{frame}

@@

    Now, when two clients issue concurrent deposit transactions, the two
    replicas execute the transactions immediately without consulting one
    another. And every once in a while, the two replicas communicate and fill
    each other in on the transactions they missed.

    Here, for example, replica A lets replica B know to deposit 10 dollars and
    replica B lets replica A know to deposit 20 dollars. Notice that even
    though we didn't perform any coordination, the two replicas never violate
    my critical invariant of a non-negative balance.
  }
\end{frame}

>>>>

  \note{%
    One way to replicate our database is by using coordination to implement
    strong consistency. For example, imagine a user issues a deposit of 10
    dollars to replica A, and simultaneously, a user issues a deposit of 20
    dollars to replica B. The two replicas can coordinate, for example by
    running Paxos, and agree on a single total order in which to execute the
    two transactions. Here, the replicas agree to deposit 10 dollars and then
    deposit 20 dollars. Both transactions succeed.
  }
\end{frame}

@@

  \note{%
    Now, assume two clients issue concurrent withdrawal transactions of 30
    dollars and 40 dollars. Again, the two replicas coordinate and agree an a
    total order in which to execute transactions.

    They first withdraw 30 dollars, bringing my bank account balance to 20
    dollars. They then try to withdraw of 40 dollars but abort the transaction
    because I only have 40 dollars left. And while the transaction failed,
    you'll note that we never violated our application invariant.
  }
\end{frame}

@@

    Now, when two clients issue concurrent deposit transactions, the two
    replicas execute the transactions immediately without consulting one
    another. And every once in a while, the two replicas communicate and fill
    each other in on the transactions they missed. Notice that even though we
    didn't perform any coordination, the two replicas never violate my critical
    invariant of a non-negative balance.
  }
\end{frame}

### 2019-08-18 doc/vldb_slides/sections/evaluation.tex
More slide tweaks.
Slimmed down talk.

    This code specifies the running example that we've been looking at
    throughout the talk. $x$ is an integer merged by max that is initially 0.
    $y$ is an integer merged by max that is initially 0. We have a transaction
    incx that increments x and a function decy that decrements y. And, our
    invariant is that the product of $x$ and $y$ is less than or equal to zero.
  }
\end{frame}

@@

  \note{%
    We have some graphs showing that no coordination vastly outperforms
    replicating with strong consistency, and how segmented invariant confluence
    allows a smooth tradeoff between the two.
  }
\end{frame}

>>>>

    % This code specifies the running example that we've been looking at
    % throughout the talk. $x$ is an integer merged by max that is initially 0.
    % $y$ is an integer merged by max that is initially 0. We have a transaction
    % incx that increments x and a function decy that decrements y. And, our
    % invariant is that the product of $x$ and $y$ is less than or equal to zero.
  }
\end{frame}

@@

  \note{%
    We have some graphs showing that no coordination vastly outperforms
    replicating with strong consistency, and how segmented invariant confluence
    allows a smooth tradeoff between the two.

    % Fully invariant confluent objects don't require any coordination, partly
    % invariant confluent objects require some coordination, and not at all
    % invariant confluent objects require a lot of coordination.
  }
\end{frame}

### 2019-08-18 doc/vldb_slides/sections/iclosure.tex
Slimmed down talk.

\begin{frame}
  \Large
  We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
  y) \in S$. \pause For example,
  \begin{itemize}
    \item Even numbers are closed under addition.
    \item Odd numbers are \emph{not} closed under addition.
  \end{itemize}

  \note{%
    First, a quick refresher on what it meas for a set to be closed. We say a
    set $S$ is closed under a binary operator $f$ if for every $x$ and $y$ in
    $S$, $f(x, y)$ is also in $S$. For example, even numbers are closed under
    addition because the sum of any two even numbers is even. But odd numbers
    are not closed under addition because the sum of two odd numbers may not
    be odd.
  }
\end{frame}

@@

  \note{%
    Invariant closure is a sufficient condition for invariant confluence. Why
    is that? Well for an object to be invariant confluent, every reachable
    state must satisfy the invariant. A state is reachable if it can be reached
    by executing transactions merging states. Transactions are already
    guarnateed to preserve invariants, so if merging is also guaranteed to
    preserve the invariant, then every reachable state always satisfies the
    invariant.
  }
\end{frame}

@@

  \note{%
    The good thing about invariant closure is that it's much easier to check
    automatically. Remember that example we had with pairs of
    integers? We can pose whether or not that object is invariant closed as
    this formula, and we can pass this formula directly to an SMT solver to
    figure out whether it's invariant closed. This is not something we can do
    with invariant confluence because it's hard to express reachability as a
    simple formula like this one.
  }
\end{frame}

@@

  \note{%
    To decide whether an object is invariant confluent, we can take the
    object and ask if it's invariant closed. If it is, then it's also invariant
    confluent and we're done. If it's not invariant closed, then what do we
    know?
  }
\end{frame}

@@

    Why? Well, the forward direction here is what we proved before. For the
    backwards direction, we have that the invariant is a subset of the set of
    reachable points, and because our object is invariant confluent, the set of
    reachable points is a subset of the invariant. So the invariant and the set
    of reachable points are the same. The set of reachable states is closed
    under merge, so the invariant is also, so the object is invariant closed.
    \\[12pt]

    I know I've walked through a lot of theory pretty quickly, but if you
    remember one slide, remember this one. If the invariant is a subset of the
    reachable states, then invariant closure and invariant confluence are
    equivalent. \\[12pt]
  }
\end{frame}

>>>>

% \begin{frame}
%   \Large
%   We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
%   y) \in S$. \pause For example,
%   \begin{itemize}
%     \item Even numbers are closed under addition.
%     \item Odd numbers are \emph{not} closed under addition.
%   \end{itemize}
%
%   \note{%
%     First, a quick refresher on what it meas for a set to be closed. We say a
%     set $S$ is closed under a binary operator $f$ if for every $x$ and $y$ in
%     $S$, $f(x, y)$ is also in $S$. For example, even numbers are closed under
%     addition because the sum of any two even numbers is even. But odd numbers
%     are not closed under addition because the sum of two odd numbers may not
%     be odd.
%   }
% \end{frame}

@@

  \note{%
    Invariant closure is a sufficient condition for invariant confluence. I'll
    refer you to the paper for the proof, but it's a very simple proof.

    % Why is that? Well for an object to be invariant confluent, every reachable
    % state must satisfy the invariant. A state is reachable if it can be reached
    % by executing transactions merging states. Transactions are already
    % guarnateed to preserve invariants, so if merging is also guaranteed to
    % preserve the invariant, then every reachable state always satisfies the
    % invariant.
  }
\end{frame}

@@

  \note{%
    The good thing about invariant closure is that it's much easier to check
    automatically. Remember that example we had with pairs of
    integers? We can pose whether or not that object is invariant closed as
    this formula, and we can pass this formula directly to an SMT solver to
    figure out whether it's invariant closed.

    % This is not something we can do with invariant confluence because it's hard
    % to express reachability as a simple formula like this one.
  }
\end{frame}

@@

  \note{%
    To decide whether an object is invariant confluent, then, we can take the
    object and ask if it's invariant closed. If it is, then it's also invariant
    confluent and we're done. If it's not invariant closed, then what do we
    know?
  }
\end{frame}

@@

    Again, I'll refer you to the paper for a proof.
    % Why? Well, the forward direction here is what we proved before. For the
    % backwards direction, we have that the invariant is a subset of the set of
    % reachable points, and because our object is invariant confluent, the set of
    % reachable points is a subset of the invariant. So the invariant and the set
    % of reachable points are the same. The set of reachable states is closed
    % under merge, so the invariant is also, so the object is invariant closed.
    % \\[12pt]
    %
    % I know I've walked through a lot of theory pretty quickly, but if you
    % remember one slide, remember this one. If the invariant is a subset of the
    % reachable states, then invariant closure and invariant confluence are
    % equivalent. \\[12pt]
  }
\end{frame}
